Project-I by Group Sydney

### 2015-11-02 report/paper.tex
Add final submission

\begin{figure}[!ht]
\vspace{-2em}
\center
\subfigure[Histogram of $\mathbf{y}$. We can see three Gaussian distributions and also some outliers on the right.]{\includegraphics[width=2.7in]{figures/histogram.jpg} \label{fig:histogram}}
\hfill
\subfigure[Feature 2 can help us to separate the data. Green data points are misclassified data. The separation is at $x=0.42$.]{\includegraphics[width=2.7in]{figures/feature2.jpg}\label{fig:feature2}}
\hfill
\subfigure[Feature 16 can help us to separate the data. Green data points are misclassified data. The separation is at $x=1.17$.]{\includegraphics[width=2.7in]{figures/feature16.jpg}\label{fig:feature16}}
\hfill
\subfigure[Different degree for the polynomial basis of the cluster 1. We can observe a low variance. The third degree seems the best model to choose.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
\vspace{-1em}
\caption{}
\end{figure}

@@

We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
\vspace{-0.5em}
\subsection{Ridge Regression}

@@

At this stage, we want to try to separate our data into clusters. The fourth model consists of using a constant model but this time separating the data into three clusters and then computing the mean for each cluster. We can observe a big improvement, justifying our choice to split the data into clusters. The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda. We also use $k$ fold cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k. As expected we obtained another big improvement justifying again the choice of splitting the data.

@@

The two next models is about using polynomial basis functions for each cluster. Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.  We have to be careful about the polynomial degrees to avoid the case where $D > N$. Moreover, if we don't observe significant improvement between degrees, we keep the lowest one. The Figure \ref{fig:degrees_cls1}, shows the case for the first cluster, we can also see that the variance is very low (same observation for the third one, however the second one has higher variance). The seventh model doesn't use dummy encoding and the eighth does.

@@

We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was too slow and less efficient than ridge regression. Separation of the data into clusters was essential in order to decrease the test error. Unfortunately, dummy encoding didn't help. However using polynomial basis helped to decrease the bias. Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher. We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point. The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.34$ with a standard deviation of $454.61$. However, we should take into consideration outliers, which gives us a \textit{RMSE} of $412.52$ with a standard deviation of $234.78$. Finally, we think that our model is improvable, however we couldn't try other things to improve the estimation error, due to a lack of time.

>>>>

\begin{figure}[!ht]
\vspace{-2em}
\center
\subfigure[Histogram of $\mathbf{y}$. We can see three Gaussian distributions and also some outliers on the right.]{\includegraphics[width=2.7in]{figures/histogram.jpg} \label{fig:histogram}}
\hfill
\subfigure[Feature 2 can help us to separate the data. Light blue data points are misclassified data. The separation is at $x=0.42$.]{\includegraphics[width=2.7in]{figures/feature2.jpg}\label{fig:feature2}}
\hfill
\subfigure[Feature 16 can help us to separate the data. Light blue data points are misclassified data. The separation is at $x=1.17$.]{\includegraphics[width=2.7in]{figures/feature16.jpg}\label{fig:feature16}}
\hfill
\subfigure[Different degree for the polynomial basis of the cluster 1. We can observe a low variance. The third degree seems the best model to choose.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
\vspace{-1em}
\caption{}
\end{figure}

@@

We can note that our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
\vspace{-0.5em}
\subsection{Ridge Regression}

@@

At this stage, we want to try to separate our data into clusters. The fourth model consists of using a constant model but this time separating the data into three clusters and then computing the mean for each cluster. We can observe a big improvement, justifying our choice to split the data into clusters. The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda. We also use $k$ fold cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really have a small number of data ($273$) and so we have to choose a smaller k. As expected we obtained another big improvement justifying again the choice of splitting the data.

@@

The two next models are about using polynomial basis functions for each cluster. Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.  We have to be careful about the polynomial degrees to avoid the case where $D > N$. Moreover, if we don't observe significant improvement between degrees, we keep the lowest one. The Figure \ref{fig:degrees_cls1}, shows the case for the first cluster, we can also see that the variance is very low (same observation for the third one, however the second one has higher variance). The seventh model doesn't use dummy encoding and the eighth does.

@@

We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was too slow and less efficient than ridge regression. Separation of the data into clusters was essential in order to decrease the test error. Unfortunately, dummy encoding didn't help. However using polynomial basis helped to decrease the bias. Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher. We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point. The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.34$ with a standard deviation of $44.62$. However, we should take into consideration outliers, which gives us a \textit{RMSE} of $412.52$ with a standard deviation of $234.78$. Finally, we think that our model is improvable, however we couldn't try other things to improve the estimation error, due to a lack of time.
